---
title: "Regression Project"
author: "Adrien Toulouse"
date: "11/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(corrplot)
library(ggplot2)
```

```{r}
trainImputed <- read.csv(file = 'train_imputed.csv')
trainPP <- read.csv(file = 'train_preprocessed.csv')
```

```{r}
#str(trainImputed)
```

```{r}
#summary(trainImputed)
```

#1. Log transformations 

We reduce the effect of the tails in the distribution of the variables using log transformations. We transform 4 variables: SalePrice, LotArea, TotalBsmtSF, GrLivArea.

```{r}
par(mfrow=c(2,1))
plot(density(trainImputed$SalePrice))
trainImputed$LogSalePrice <- log(trainImputed$SalePrice)
trainImputed <- select(trainImputed, -c("SalePrice"))
plot(density(trainImputed$LogSalePrice))
```


```{r}
par(mfrow=c(2,1))
plot(density(trainImputed$LotArea))
trainImputed$LogLotArea <- log(trainImputed$LotArea)
trainImputed <- select(trainImputed, -c("LotArea"))
plot(density(trainImputed$LogLotArea))
```

```{r}
par(mfrow=c(2,1))
plot(density(trainImputed$TotalBsmtSF))
trainImputed$LogTotalBsmtSF <- log(trainImputed$TotalBsmtSF)
trainImputed <- select(trainImputed, -c("TotalBsmtSF"))
plot(density(trainImputed$LogTotalBsmtSF))
```

```{r}
par(mfrow=c(2,1))
plot(density(trainImputed$GrLivArea))
trainImputed$LogGrLivArea <- log(trainImputed$GrLivArea)
trainImputed <- select(trainImputed, -c("GrLivArea"))
plot(density(trainImputed$LogGrLivArea))
```


# 2. Numeric variables: 

Looking at the numeric variables, we check correlation between the different variables together and then the correlation of SalePrice with the variables.
The first step is not useful when you try to predict a variable. It is however very important when you try to see which variables are the most important to explain price. (Remember the p-value of statistical test for correlated variables). 

```{r}
var.numeric <- colnames(trainPP)[sapply(trainPP, is.numeric)]

trainPP %>%
  select(var.numeric) %>%
  cor() %>%
  corrplot(method = 'color', order = "hclust", tl.pos = 'n') %>%
  heatmap (symm=T)
```

```{r}
ggplot(trainImputed, aes(LotFrontage, LogLotArea)) + geom_point() + geom_smooth(method=lm)
```

```{r}
ggplot(trainImputed, aes(LogTotalBsmtSF, Fireplaces)) + geom_point() + geom_smooth(method=lm)
```

```{r}
ggplot(trainImputed, aes(YearBuilt, YearRemodAdd)) + geom_point() + geom_smooth(method=lm)
```

```{r, message=FALSE, warning=FALSE}
trainPP %>%
  select(var.numeric) %>%
  GGally::ggpairs(columns = c(length(var.numeric), 1:5))
```

#Let's now look at the correlations with our response variables.

```{r}
ggplot(trainImputed, aes(OverallQual, LogSalePrice)) + geom_point() + geom_smooth(method=lm)
```

Overall quality impact Sale Price. Looks like a linear relationship. we can include this predictor without doing any transformation. 
```{r}
ggplot(trainImputed, aes(OverallCond, LogSalePrice)) + geom_point()
```

```{r}
ggplot(trainImputed, aes(LogTotalBsmtSF, LogSalePrice)) + geom_point() + geom_smooth(method=lm)
```

```{r}
ggplot(trainImputed, aes(LogLotArea, LogSalePrice)) + geom_point() + geom_smooth(method=lm)
```
```{r}
ggplot(trainImputed, aes(LotFrontage, LogSalePrice)) + geom_point() + geom_smooth(method=lm)
```

```{r}
ggplot(trainImputed, aes(YearBuilt, LogSalePrice)) + geom_point() + geom_smooth(method=lm)
```
```{r}
ggplot(trainImputed, aes(YearRemodAdd, LogSalePrice)) + geom_point() + geom_smooth(method=lm)
```

```{r}
ggplot(trainImputed, aes(Fireplaces, LogSalePrice)) + geom_point() + geom_smooth(method=lm)
```

#3. Factor variables: 

We plot the SalePrice as a factor of the factor variables by using boxplots. 

```{r}
ggplot(data = trainImputed) + 
  geom_boxplot(aes(y=LogSalePrice, x = CentralAir))
```

```{r}
ggplot(data = trainImputed) + 
  geom_boxplot(aes(y=LogSalePrice, x = RoofMatl))
```

```{r}
ggplot(data = trainImputed) + 
  geom_boxplot(aes(y=LogSalePrice, x = GarageQual))
```

May have an impact but doesn't follow totally intuition. 

```{r}
ggplot(data = trainImputed) + 
  geom_boxplot(aes(y=LogSalePrice, x = MSZoning))
```

Looks like it is pertinent. Let's do a statistical test to show that there is a correlation statistically significant between the Sale price and the MSZoning.
```{r}
res.aov <- aov(LogSalePrice ~ MSZoning, data = trainImputed)
summary(res.aov)
```
As the p-value is less than the significance level 0.05, we can conclude that there are significant differences between the different MSZoning and make us say that we should include MSZoning in the model. 

```{r}
ggplot(data = trainImputed) + 
  geom_boxplot(aes(y=LogSalePrice, x = LotConfig))
```

```{r}
res.aov <- aov(LogSalePrice ~ LotConfig, data = trainImputed)
summary(res.aov)
```


```{r}
ggplot(data = trainImputed) + 
  geom_boxplot(aes(y=LogSalePrice, x = Neighborhood))
```

#4. Ancova analysis

Here we try to plot two quantitative variables together with one factor variable. 

```{r}
p <- ggplot(trainImputed, aes(y=LogSalePrice, x=LogGrLivArea, color = MSZoning, fill = MSZoning)) + 
  geom_point() + theme_classic()
p + stat_smooth(method = "lm") 
```

This helps to make clear that while GrLiveArea has a large predictive effect for LogSalePrice (the slopes of all the lines are clearly non-zero), there is also an effect of group assignment: for example the houses assigned to the FV MSZoning have a higher Sale Price than the houses assigned to RH.  

```{r}
p <- ggplot(trainImputed, aes(y=LogSalePrice, x=LogTotalBsmtSF, color = CentralAir, fill = CentralAir)) + 
  geom_point() + theme_classic()
p + stat_smooth(method = "lm") 
```

This shows that while TotalBsmtSF has a large predictive effect for LogSalePrice (the slopes of the two lines are clearly non-zero), there is also an effect of group assignment: the houses with CentralAir have a higher LogSale Price. However since the slopes are the same, there seems to be no correlation between having CentralAir and the TotalBsmtSF. 



```{r}
p <- ggplot(trainImputed, aes(y=LogSalePrice, x=LogLotArea, color = MSZoning, fill = MSZoning)) + 
  geom_point() + theme_classic()
p + stat_smooth(method = "lm") 
```


MSZoning + LotArea + OverallQual + OverallCond + YearBuilt + YearRemodAdd + RoofMatl + TotalBsmtSF + CentralAir + GrLivArea + KitchenQual + Fireplaces + GarageQual